---
title: "Optimisation"
author: "Vishruth Reddy"
date: "2022-12-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After you have selected the best performing models consider:
• Does the model performance improve when the derived features in the “expanded feature” set are included?
yes it does.
• Identify the most important variables associated with your best performing models.
z, w, t, x5 
• Visualize the predicted logit-transformed response as a function of your identified most important
variables.
• Visualize the predicted probability of the EVENT as a function of your identified most important variables.
• Based on your visualizations, what input settings are associated with minimizing the logit-transformed response?
• Do the optimal input settings vary across the values of the categorical variable?
• BONUS +10 points: Optimize the inputs/features for 2 values of the categorical variable using optim(). 

```{r, read_final_data}
df <- readr::read_csv("fall2022_finalproject.csv", col_names = TRUE)
```

```{r}
library(dplyr)
```


```{r, show_data_glimpse}
df %>% glimpse()
```


```{r, show_derived_features}
df <- df %>% 
  mutate(x5 = 1 - (x1 + x2 + x3 + x4),
         w = x2 / (x3 + x4),
         z = (x1 + x2) / (x5 + x4),
         t = v1 * v2)
```


```{r, show_logit_transform}
df <- df %>% 
  mutate(y = boot::logit(output)) 
```

```{r, show_binary_outcome}
df <- df %>% 
  mutate(outcome = ifelse(output < 0.33, 'event', 'non_event'),
         outcome = factor(outcome, levels = c("event", "non_event"))) 
```



```{r}
df
```

```{r}
library(caret)
```

```{r}
mars_expanded <- readr::read_rds("mars_reg.rds")
xgb_expanded_glm <- readr::read_rds("xgb_class.rds")
```

```{r}
ctrl_roc <- trainControl( method = "repeatedcv", number = 5, repeats = 3,
                         summaryFunction = twoClassSummary,
                         classProbs = TRUE,
                         savePredictions = TRUE)

```

```{r}
marsGrid <- expand.grid(
  degree = 1:3, 
  nprune = seq(2, 100, length.out = 10) %>% floor()
  )
```

```{r}
mars_expanded_roc<- caret::train(outcome ~ x1  + x3 + x4  + v2 + v3 + v4 + v5 + m + w + z + t + x5,
                            data = df,
                            method = "earth", 
                            tuneGrid = marsGrid,
                            preProcess = c("center", "scale"),
                            metric = "ROC",
                            trControl = ctrl_roc)

```

```{r}
mars_expanded_roc
```

```{r}
model_reg <- mars_expanded
model_reg
```


```{r}
model_class <- xgb_expanded_glm
model_class
```

```{r}
reg_preds <- predict(model_reg, df)
reg_preds
```


```{r}
dim(df)
dim(reg_preds)
```

```{r}
df['probs'] <- reg_preds
```

```{r}
df
```


```{r}
df %>% ggplot(mapping=aes(x = x5, y = probs)) + geom_point()
df %>% ggplot(mapping=aes(x = w, y = probs)) + geom_point()
df %>% ggplot(mapping=aes(x = z, y = probs)) + geom_point()
df %>% ggplot(mapping=aes(x = t, y = probs)) + geom_point()

```



```{r}
class_preds <- predict(model_class, df, type = 'prob')
class_preds
```


```{r}
df['event'] <- class_preds$event
df['non_event'] <- class_preds$non_event
```

```{r}
df
```





```{r}
df %>% ggplot(mapping=aes(x = x5, y = event)) + geom_point()
df %>% ggplot(mapping=aes(x = w, y = event)) + geom_point()
df %>% ggplot(mapping=aes(x = z, y = event)) + geom_point()
df %>% ggplot(mapping=aes(x = t, y = event)) + geom_point()

```

```{r}
test_df <- readr::read_csv("fall2022_holdout_inputs.csv", col_names = TRUE)

```


```{r}
test_df

```



```{r}
test_df_ <- test_df %>% 
  mutate(x5 = 1 - (x1 + x2 + x3 + x4),
         w = x2 / (x3 + x4),
         z = (x1 + x2) / (x5 + x4),
         t = v1 * v2)
test_df_
```


```{r}

test_class_preds <- predict(xgb_expanded_glm, test_df_)
test_class_prob <- predict(xgb_expanded_glm, test_df_, type = 'prob')
test_class_y <- predict(xgb_expanded_glm, test_df_, type = 'raw')
```



```{r}
test_reg_preds <- predict(model_reg, test_df_)
```


```{r}
final_df <- data.frame('id' = seq(1,nrow(test_df_)), 'outcome' = test_class_preds, 'probability' =  test_class_prob$event, 'y' = test_reg_preds)
final_df

```

```{r}
write.csv(final_df, "preds.csv", row.names=FALSE)
```
